Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-09-11 scagnostics.Rmd
Working out some details for example distributions.

@@

### Looking down the pipe
TIC should be able to uncover special view, compare performance of all indices on guided tour path. (also cmass and holes?)
Not currently finding "best" view. Options: compare results from n different random starting planes, try with smaller space (e.g. omitting variable 1 and 2), turn some of the guided tour parameters
```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset."}
set.seed(2018)
pipeResc <- rescale(pipe1000)
pipeTour <- save_history(pipeResc,
                         guided_tour(mineIndex("TIC")), start = p58)
pipeTourFull <- as.list(interpolate(pipeTour))
pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
ggplot(pipeTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

>>>>

To get an overview of the dataset we show views of the first two parameters showing the independent distribution and the last two parameters showing the special view. The enforces relationships will also affect the overall distribution as can be seen from (fig:dataPlotsDensity).

```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V2."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V1, V2)) +
    geom_point() +
    ggtitle(pLabels[i])
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Data views of V5 vs V6."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V5, V6)) +
    geom_point() +
    ggtitle(pLabels[i])
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="Data views of V1 vs V6."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  distData <- melt(ds)
  pC <- ggplot(distData, aes(value, color=variable)) +
    geom_density() +
    ggtitle(pLabels[i])
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

We next compare the various indexvalues for three views: V1 vs V2, V5 vs V6 and V1 vs V6 (capturing differences from the different distribution in V6).

```{r indexTable, echo=FALSE}
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
indexOverview <- tibble(convex=numeric(),
               clumpy=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC=numeric(),
               name=character(),
               vars=character())
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  ds <- as.tibble(rescale(ds))
  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    indexOverview <- add_row(indexOverview, convex=scagRes["Convex"],
                    clumpy=scagRes["Clumpy"], skinny=scagRes["Skinny"],
                    stringy=scagRes["Stringy"], dcor2d=dcorRes,
                    splines2d=splineRes, MIC=mineRes$MIC,
                    TIC=mineRes$TIC, name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
  }
  i <- i+1
}
knitr::kable(indexOverview ,  caption = "Summary of index values.", digits=2) 
```

```

@@

### Looking down the pipe
TIC should be able to uncover special view, compare performance of all indices on guided tour path. (also cmass and holes?)
Not currently finding "best" view. Options: compare results from n different random starting planes, try with smaller space (e.g. omitting variable 1 and 2), turn some of the guided tour parameters
```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset."}
set.seed(2018)
pipeResc <- rescale(pipe1000)
pipeTour <- save_history(pipeResc,
                         guided_tour(mineIndex("TIC"), n=10, cooling = 0.8), start = p58)
pipeTourFull <- as.list(interpolate(pipeTour))
pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
ggplot(pipeTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```
